---
title: "Book Reviews"
author: "Elijah Silfies"
date: "11/22/2021"
output: html_notebook
---
In this project I will be analyzing book sales data. On July 1st, 2019, this company changed their marketing strategy, so I will be showing the affect of that new strategy on their book sales.

The variables in this data consist of:

date = The date of the purchase in 2019

user_submitted_review = The review on the book from the customer

title = The title of the book purchased

total_purchased = the quantity of books purchased by that customer

customer_type = business or individual based purchased 

```{r}
library(tidyverse)
library(lubridate)
sales2019 <- read.csv("sales2019.csv")
```

### Analyse the data
```{r}
glimpse(sales2019)
```
The user_submitted_review column and the total_purchased column are both missing some data values. Also total_purchased is in integer form, so I will be converting this to a double to save some trouble doing calculations.

### Change total-purchased column to double variable
```{r}
sales2019 <- sales2019 %>% 
  mutate(total_purchased = as.double(sales2019$total_purchased))
```

### Remove NA values in the user review column
```{r}
sales2019 <- sales2019 %>% 
  filter(!is.na(user_submitted_review))
```
A total of 885 values were lost so instead I will substitute the NA values for the mean of the total books purchased.

### Find the average of books sold per customer
```{r}
avgsales <- sales2019 %>% 
  filter(!is.na(total_purchased)) %>% 
    pull(total_purchased) %>% 
      mean

print(avgsales)
```

### Sub the NA values for the avereage of all books purchased
```{r}
sales <- sales2019 %>% 
  mutate(
    total_purchased = if_else(is.na(total_purchased),
    avgsales,
    total_purchased)
        )

head(sales)
```

### Find the unique reviews
```{r}
sales %>% 
  pull(user_submitted_review) %>% 
    unique
```

### Determining positive or negative reviews with a new column
```{r}
sales <- sales %>% 
  mutate(
    Positive_Review = 
      case_when(
        str_detect(user_submitted_review, pattern = "Awesome") == TRUE ~ TRUE,
        str_detect(user_submitted_review, pattern = "Never read") == TRUE ~ TRUE,
        str_detect(user_submitted_review, pattern = "learned") == TRUE ~ TRUE,
       TRUE ~ FALSE
              )
        )

head(sales)
```

### Changing the format on the dates
```{r}
sales <- sales %>% 
  mutate(
    date = mdy(date)
        )

head(sales)
```

### Do the sales happen before or after july 1st 2019?
```{r}
sales <- sales %>%
  mutate(
    afterjuly = 
      case_when(
        ymd(date) < mdy("7-1-19") ~ "no",
         ymd(date) >= mdy("7-1-19") ~ "yes"
               )
        )

head(sales)
```

### Sort the data set by before and after July 1st. Did the marketing change increace the number of books sold?
```{r}
sort_sales <- sales %>% 
  group_by(afterjuly) %>% 
    summarise(BooksPurchased =sum(total_purchased))

print(sort_sales)
```
the number of books sold did not increase after July 1st 2019. in fact there was a slight decline in sales.

### Sort the data based on before and after July 1st the type of consumer. Did they gaine purchases from a specific type of consumer?
```{r}
sort_sales_customer <- sales %>% 
  group_by(afterjuly,customer_type) %>% 
    summarise(BooksPurchased = sum(total_purchased))

print(sort_sales_customer)
```
We can now see that the number of books purchased by businesses increased and the number of books purchased by individuals decreased significantly. 

### Did the number of positive sales increase after july 1st?
```{r}
sort_sales_positive <- sales %>% 
  group_by(afterjuly) %>% 
    summarise(`Positive Reviews`= sum(Positive_Review))

print(sort_sales_positive)
```
There was a gain in about 30 positive reviews which is not a large number compared to 600, but it is still a slight increase. It may be an indication that they're on the right track
